% Appendix B: Webbaselets — BPMS, PLM, Ticketing

\chapter{Webbaselets: BPMS, PLM, and Ticketing}
\label{app:webbaselets}

This appendix summarizes three foundational webbaselets that commonly ship with a Spin the Web deployment. Each is delivered as a standalone webbaselet (root STWArea) that can be imported into any Webbase, sharing common conventions for navigation, security, and data.

\begin{itemize}
  \item Business Process Management System (BPMS)
  \item Presence and Leave Management (PLM)
  \item Ticketing (Service Desk)
\end{itemize}

All three follow the same patterns:
\begin{enumerate}
  \item Pages expose task-oriented UI
  \item A clear data model with well-known keys and relations
  \item Role-based visibility and action authorization
  \item Events and workflows that integrate across webbaselets
\end{enumerate}

\section{Common Design}
\label{sec:app-webbaselets-common}

\textbf{Navigation.} Each webbaselet registers under its own Area (e.g., slugs \texttt{bpms}, \texttt{plm}, \texttt{tickets}); pages include Dashboard, Browse/Detail, and Administration.

\textbf{Security.} Every action checks roles and ownership. Typical roles include: reader, contributor, approver, manager, and admin. Actions are logged.

\textbf{Search and SEO.} Content uses \texttt{keywords} and \texttt{description} attributes for in-context search. Lists expose filters and faceted navigation.

\textbf{Auditability.} All state transitions are captured with who/when/what; attachments preserved with versions.

\textbf{Interoperability.} Events are published as notifications that other webbaselets can subscribe to (e.g., Ticket created -> BPMS starts a triage workflow; Leave approved -> Ticket auto-closes).

\subsection{Minimal Area Stub}
A minimal root for each webbaselet as an STWArea:
\begin{lstlisting}[language=JSON,caption={Minimal Webbaselet Root (Area)}]
{
  "_id": "area-bpms-root",
  "type": "Area",
  "namespace": "bpms",
  "name": { "en": "BPMS" },
  "slug": { "en": "bpms" },
  "children": []
}
\end{lstlisting}

\section{Structured Classification for Webbaselets}
\label{sec:app-webbaselets-classification}

To make webbaselets discoverable and governable across the ecosystem, the root \texttt{STWArea} of each webbaselet SHOULD declare a compact, structured \emph{classification}. We recommend a dedicated \texttt{class} object on the Area to avoid collisions with core keys; engines MAY also surface these as indexed metadata for search.

\subsection{Classification keys}
Recommended fields and value ranges:

\begin{tabularx}{\linewidth}{@{}l l X@{}}
  \textbf{Field} & \textbf{Type} & \textbf{Purpose / Examples} \\
\hline
  \texttt{purpose} & enum & Primary intent: \texttt{bpms}, \texttt{plm}, \texttt{ticketing}, \texttt{cms}, \texttt{directory}, \texttt{reporting}, \texttt{integration}, \texttt{admin}. \\
  \texttt{domain} & enum & Business area: \texttt{hr}, \texttt{it}, \texttt{finance}, \texttt{sales}, \texttt{marketing}, \texttt{ops}, \texttt{legal}, \texttt{support}, \texttt{general}. \\
  \texttt{capability} & string[] & Features: \texttt{workflow}, \texttt{approvals}, \texttt{forms}, \texttt{queueing}, \texttt{sla}, \texttt{search}, \texttt{analytics}, \texttt{files}, \texttt{notifications}, \texttt{calendar}. \\
  \texttt{layer} & enum & Architectural layer: \texttt{ui}, \texttt{service}, \texttt{integration}, \texttt{data}. \\
  \texttt{audience} & string[] & Users: \texttt{employees}, \texttt{managers}, \texttt{admins}, \texttt{agents}, \texttt{external}. \\
  \texttt{criticality} & enum & \texttt{low}, \texttt{medium}, \texttt{high}, \texttt{critical}. \\
  \texttt{maturity} & enum & \texttt{experimental}, \texttt{beta}, \texttt{stable}, \texttt{deprecated}. \\
  \texttt{visibility} & enum & \texttt{public}, \texttt{authenticated}, \texttt{role-restricted}. \\
  \texttt{compliance} & string[] & \texttt{gdpr}, \texttt{hipaa}, \texttt{sox}, \texttt{pci}, \texttt{iso27001}, etc. \\
  \texttt{dataSensitivity} & enum & \texttt{public}, \texttt{internal}, \texttt{confidential}, \texttt{pii}, \texttt{phi}. \\
  \texttt{i18n} & string[] & Supported locales (BCP\,47): e.g., \texttt{["en","it"]}. \\
  \texttt{engine} & object & Runtime bounds: \texttt{{"min": "1.3.0", "max": "<2.0.0"}}. \\
  \texttt{dependsOn} & string[] & Other webbaselets by namespace and range, e.g., \texttt{["directory>=1.2", "notifications"]}. \\
  \texttt{imports} & string[] & Subscribed events/exports (namespaced), e.g., \texttt{["bpms.task.completed"]}. \\
  \texttt{exports} & string[] & Emitted events/components, e.g., \texttt{["tickets.created", "ui.widget:calendar"]}. \\
  \texttt{tags} & string[] & Free-form keywords for discovery. \\
\end{tabularx}

\subsection{Example}
BPMS classification on its root Area:
\begin{lstlisting}[language=JSON,caption={STWArea with classification}]
{
  "_id": "area-bpms-root",
  "type": "Area",
  "namespace": "bpms",
  "name": { "en": "BPMS" },
  "slug": { "en": "bpms" },
  "class": {
    "purpose": "bpms",
    "domain": "hr",
    "capability": ["workflow","approvals","forms","sla"],
    "layer": "ui",
    "audience": ["employees","managers","admins"],
    "criticality": "high",
    "maturity": "stable",
    "visibility": "authenticated",
    "compliance": ["gdpr"],
    "dataSensitivity": "pii",
    "i18n": ["en","it"],
    "engine": { "min": "1.3.0" },
    "dependsOn": ["directory>=1.1", "notifications"],
    "imports": ["tickets.created"],
    "exports": ["bpms.task.created","bpms.task.completed"],
    "tags": ["process","tasks","orchestration"]
  },
  "children": []
}
\end{lstlisting}

\paragraph{Notes}
\begin{itemize}
  \item Engines should index \texttt{class.*} for search and dependency checks.
  \item Recommending enums keeps catalogs consistent; use \texttt{tags} for additional nuance.
  \item Version ranges follow semantic versioning; omit \texttt{max} for open-ended support.
\end{itemize}

\section{Ecosystem Webbaselet (Community Catalog)}
\label{sec:app-ecosystem}

The Spin the Web portal (see \cref{part:implementation}) includes an \emph{Ecosystem} webbaselet that catalogs community-created webbaselets. Its purpose is to classify, make discoverable, and enable quality signals (rating/reviews) for webbaselets across the ecosystem. It builds directly on the structured classification in \cref{sec:app-webbaselets-classification} and extends it with an opinionated, organization-based taxonomy.

\subsection{Organization-based taxonomy}
Classification aligns to common enterprise departments to aid discovery by function:
\begin{tabularx}{\linewidth}{@{}l X@{}}
  \textbf{Domain} & \textbf{Examples} \\
\hline
hr & onboarding, leave, compensation, training \\
it & service desk, change, assets, access management \\
finance & expenses, invoicing, budgeting, approvals \\
sales & leads, opportunities, quotes, forecasts \\
marketing & campaigns, content hub, events, assets \\
ops & facilities, fleet, logistics, maintenance \\
legal & contracts, policies, compliance \\
procurement & vendors, sourcing, purchase requests \\
support & knowledge base, ticket deflection, CSAT \\
rnd & requirements, roadmaps, experiments \\
product & releases, feature flags, feedback \\
security & incidents, awareness, audits \\
admin & directory, identity, governance \\
general & dashboards, search, navigation, utilities \\
\end{tabularx}

These values map to \texttt{class.domain} and inform facets in catalog pages.

\subsection{Data model}
Key entities managed by the Ecosystem webbaselet:
\begin{lstlisting}[language=JSON,caption={Ecosystem entities}]
{
  "catalogItem": {
    "_id": "eco-wbl-hr-directory",
    "namespace": "hr-directory",
    "name": {"en":"HR People Directory"},
    "summary": {"en": "Org-wide searchable people directory"},
    "class": {
      "purpose": "directory",
      "domain": "hr",
      "capability": ["search","profiles"],
      "tags": ["people","org","contact"]
    },
    "version": "1.2.0",
    "author": {"name":"Community Team","url":"https://spintheweb.org"},
    "links": {"repo":"https://...","docs":"https://...","demo":"https://..."},
    "status": "published"
  },
  "review": {
    "_id": "rev-123",
    "item": "eco-wbl-hr-directory",
    "user": "u:alice",
    "stars": 5,
    "comment": "Great starter for HR portals",
    "createdAt": "2025-09-12T10:00:00Z"
  },
  "rating": { "item": "eco-wbl-hr-directory", "average": 4.6, "count": 28 }
}
\end{lstlisting}

\subsection{Workflows}
\begin{itemize}
  \item Submission: contributor provides metadata, classification, and links; item enters \emph{submitted} state.
  \item Review/Moderation: approvers validate classification, security notes, and licensing; states: \emph{approved}, \emph{rejected}, \emph{changes-requested}.
  \item Publish/Deprecate: approved items are listed; subsequent versions can supersede; deprecated items remain searchable with badges.
  \item Rating/Reporting: authenticated users rate (1--5) and review; abuse reports trigger moderation.
\end{itemize}

\subsection{Representative pages}
\begin{itemize}
  \item Catalog: faceted browse by \texttt{domain}, \texttt{purpose}, \texttt{capability}, and tags; sort by rating and freshness.
  \item Item Detail: metadata, installation notes, screenshots, version history, ratings, related items.
  \item Submit/Manage: guided submission, preview, and moderation queue (admin).
\end{itemize}

\subsection{Ratings and quality signals}
\begin{itemize}
  \item 1--5 star ratings with weighted average; first-party installs may carry a \emph{verified} badge.
  \item Anti-spam: one rating per account per version; optional cool-down windows; abuse reporting.
  \item Signals: popularity (installs), freshness (recent updates), completeness (docs, tests), compliance tags.
\end{itemize}

\subsection{Integration}
\begin{itemize}
  \item Events: \texttt{ecosystem.item.published}, \texttt{ecosystem.item.updated}, \texttt{ecosystem.review.created}.
  \item Dependencies: Directory (profiles), Notifications, Search/Index.
  \item Portal linkage: featured items and curated collections surfaced on spintheweb.org (see \cref{part:implementation}).
\end{itemize}

\section{BPMS Webbaselet}
\label{sec:app-bpms}

Purpose: model, execute, and monitor business processes with tasks, SLAs, and approvals.

\subsection{Core Concepts}
\begin{description}
  \item[Process] Named workflow definition with versioning
  \item[Instance] Runtime execution of a Process
  \item[Task] Unit of work assigned to a role/user
  \item[Event] Signal that triggers transitions or external actions
\end{description}

\subsection{Data Model}
\begin{lstlisting}[language=JSON,caption={BPMS entities}]
{
  "process": { "_id": "proc-id", "name": {"en": "Onboarding"}, "version": 3, "stages": ["init","docs","it","done"] },
  "instance": { "_id": "pi-123", "process": "proc-id", "stage": "docs", "state": "active", "assignees": ["hr:manager"], "due": "2025-09-30" },
  "task": { "_id": "task-789", "instance": "pi-123", "type": "approve", "owner": "hr:manager", "status": "open", "slaHours": 48 },
  "event": { "kind": "task.completed", "ref": "task-789", "by": "hr:manager", "at": "2025-09-12T08:30:00Z" }
}
\end{lstlisting}

\subsection{Representative Pages}
\begin{itemize}
  \item Dashboard: My Tasks, Overdue, SLA breaches
  \item Process Designer: stage/transition editor (admin)
  \item Instance Detail: timeline, attachments, comments
\end{itemize}

\subsection{WBLL Snippet}
\begin{lstlisting}[language=WBLL,caption={My Tasks list (sketch)}]
\s('caption="My Tasks" key="taskId"')
lf l f l f // id, type, status
\end{lstlisting}

\section{PLM Webbaselet}
\label{sec:app-plm}

Purpose: track presence, leave requests, balances, holidays, and approvals.

\subsection{Data Model}
\begin{lstlisting}[language=JSON,caption={PLM entities}]
{
  "calendar": { "year": 2025, "holidays": ["2025-01-01","2025-12-25"] },
  "balance": { "user": "u:alice", "year": 2025, "vacation": 15, "sick": 5 },
  "leave": { "_id": "lv-42", "user": "u:alice", "from": "2025-09-15", "to": "2025-09-20", "type": "vacation", "status": "pending", "approver": "mgr:bob" }
}
\end{lstlisting}

\subsection{Representative Pages}
\begin{itemize}
  \item My Calendar: presence, requests, approvals
  \item Team Calendar: manager view, conflicts
  \item Balances: accrual, consumption, carryover
\end{itemize}

\subsection{Events}
Examples: \texttt{leave.requested}, \texttt{leave.approved}, \texttt{leave.rejected}. Ticketing can subscribe to create handover tickets; BPMS can trigger onboarding/offboarding.

\section{Ticketing Webbaselet}
\label{sec:app-ticketing}

Purpose: log, track, and resolve issues/requests with queues, priorities, and SLAs.

\subsection{Data Model}
\begin{lstlisting}[language=JSON,caption={Ticket entities}]
{
  "ticket": { "_id": "t-1001", "queue": "it", "kind": "incident", "priority": "high", "status": "open", "subject": "Laptop fan noisy", "requester": "u:carol", "assignee": "it:agent01" },
  "comment": { "_id": "c-1", "ticket": "t-1001", "by": "u:carol", "when": "2025-09-12T09:00:00Z", "text": "Happens after 10min" },
  "sla": { "queue": "it", "kind": "incident", "responseHours": 4, "resolutionHours": 48 }
}
\end{lstlisting}

\subsection{Representative Pages}
\begin{itemize}
  \item Submit Ticket: form with WBLL inputs, attachments
  \item My Tickets: list, filters, bulk actions
  \item Agent Console: triage, assign, merge, close, macros
\end{itemize}

\subsection{Cross-Webbaselet Integrations}
\begin{itemize}
  \item BPMS: auto-create approval tasks for change requests
  \item PLM: auto-snooze tickets when requester is on leave
  \item Directory: enrich requester profile and permissions
\end{itemize}

\section{Roles and Permissions}
\begin{tabularx}{\linewidth}{@{}l X@{}}
\textbf{Role} & \textbf{Capabilities} \\
\hline
reader & browse lists and details \\
contributor & create/update own domain entities \\
approver & review/approve pending items \\
manager & oversee queues/teams, reports \\
admin & configure processes, queues, calendars \\
\end{tabularx}

\section{Notes on Deployment}
Each webbaselet can be versioned and deployed independently. Use namespace scoping, feature flags, and migration scripts for data evolution. Provide sample datasets for demos and QA.
